The Two Dualities of Computation: Negative and Fractional Types

### 2012-02-28 rational.tex
para in intro
diagrams

\begin{abstract}
  Every functional programmer knows about sum and product types, {{a+b}} and
  {{a*b}} respectively. Negative and fractional types, $a-b$ and $a/b$
  respectively, are much less known and their computational interpretation is
  unfamiliar and often complicated. We show that in a programming model in
  which information is preserved (such as the model introduced in our recent
  paper on \emph{Information Effects}), these types have particularly
  intuitive and natural computational interpretations. Intuitively, values of
  negative types are values that flow ``backwards'' to satisfy demands and
  values of fractional types are values that impose constraints on their
  context.  The combination of these negative and fractional types enables
  greater flexibility in programming by breaking global invariants into local
  ones that can be autonomously satisfied by a subcomputation. Furthermore,
  they allow a programmer to express a wide range of programming idioms,
  including higher-order functions, delimited continuations, speculative
  computation, logic programming, and more.
\end{abstract}

@@

@@

>>>>

\begin{abstract}

  Every functional programmer knows about sum and product types, {{a+b}} and
  {{a*b}} respectively. Negative and fractional types, $a-b$ and $a/b$
  respectively, are much less known and their computational interpretation is
  unfamiliar and often complicated. We show that in a programming model in
  which information is preserved (such as the model introduced in our recent
  paper on \emph{Information Effects}), these types have particularly
  intuitive and natural computational interpretations. Intuitively, values of
  negative types are values that flow ``backwards'' to satisfy demands and
  values of fractional types are values that impose constraints on their
  context.  The combination of these negative and fractional types enables
  greater flexibility in programming by breaking global invariants into local
  ones that can be autonomously satisfied by a subcomputation. Furthermore,
  they allow a programmer to express a wide range of programming idioms,
  including higher-order functions, delimited continuations, speculative
  computation, logic programming, and more.

  %% Every functional programmer knows about sum and product types,
  %% {{a+b}} and {{a*b}} respectively. Negative and fractional types,
  %% $a-b$ and $a/b$ respectively, are much less known and their
  %% computational interpretation is unfamiliar and often complicated. We
  %% show that in a programming model in which information is preserved
  %% (such as the model introduced in our recent paper on
  %% \emph{Information Effects}), these types have particularly intuitive
  %% and natural computational interpretations. Intuitively, values of
  %% negative types are values that flow ``backwards'' to satisfy
  %% demands, values of fractional types are values that represent
  %% first-class ``structural'' constraints.  The combination of these
  %% negative and fractional types enable greater flexibility in
  %% programming by breaking complicated invariants into local ones that
  %% can be independently satisfied by a subcomputation. Furthermore,
  %% they allow a programmer to express a wide range of programming
  %% idioms, including higher-order functions, delimited continuations,
  %% speculative computation, logic programming, and more.

\end{abstract}

@@

In 1989, \cite{Filinski:1989:DCI:648332.755574} observed that values and
continuations are dual notions. This observation was followed by numerous
papers (\cite{Griffin:1989:FNC:96709.96714, curien00duality, Wadler:2003,
  DBLP:conf/rta/Wadler05}) which explored and refined this duality.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Negative Types: Intuition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Fractional Types: Intuition}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Summary}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{ {{langRev}} }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Semantics}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{ {{langRevEE}} }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Additive Duality in {{langRev}} }

\begin{multicols}{2}
\begin{center}
\scalebox{1.5}{
  \includegraphics{diagrams/eps_plus1.pdf}
}
\end{center}

\begin{center}
\scalebox{1.5}{
  \includegraphics{diagrams/eps_plus2.pdf}
}
\end{center}  
\end{multicols}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Multiplicative Duality in {{langRev}} }

\begin{multicols}{2}
\begin{center}
  \includegraphics{diagrams/eta_times.pdf}
\end{center}

\begin{center}
  \includegraphics{diagrams/eps_times.pdf}
\end{center}
\end{multicols}

test test test
\begin{multicols}{2}
\begin{center}
\scalebox{1.5}{
  \includegraphics{diagrams/eta_times1.pdf}
}
\end{center}

\begin{center}
\scalebox{1.5}{
  \includegraphics{diagrams/eps_times1.pdf}
}
\end{center}  
\end{multicols}

\subsection{Syntax}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Rules for {{eta}} and {{eps}} }

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Constructions in {{langRevEE}} }
\label{sec:monoidal-constructions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Constructions Specific to Negation}
\label{sec:neg-constructions}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Constructions Specific to Multiplication}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{To think about}

\subsection{SAT solver}

\subsection{GoI machine}
